---
title: "Fantaxtic"
author: "Guus Martijn Teunisse"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fantaxtic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
##Introduction
This is fantaxtic, a set of wrapper functions for phyloseq and ggplot2 that turns ordinary taxonomic count data into fantaxtic, publication ready plots. Whereas most packages promise to deliver figures that could be published directly, in practice they often are standard plots output by ggplot2 using the minimal theme. Fantaxtic figures, however, come as completely stylized figures with pleasing colors and maximal plotting control. As the output figures are simply ggplot2 objects, expert users can also manipulate the final output to customize figures up to their desire.

##Example: Barplot of relative OTU/ASV abundances
First load the data.
```{r, include = T, fig.align = "center", fig.width = 8}
library(fantaxtic)
data(GlobalPatterns)
```

Now subset the data to the 10 most abundants OTUs or ASVs using the `get_top_taxa` function.
```{r, include = T, fig.align = "center", fig.width = 8}

ps_tmp <- get_top_taxa(physeq_obj = GlobalPatterns, n = 10, relative = TRUE,
                       discard_other = FALSE, other_label = "Other")
```

Next, create labels for missing taxonomic ranks using the `name_taxa` function.
```{r, include = T, fig.align = "center", fig.width = 8}
ps_tmp <- name_taxa(ps_tmp, label = "Unkown", species = T, other_label = "Other")
```

You are now ready to generate a `fantaxtix_bar`! Generate a plot that is colored by Phylum and labeled by Species, coloring the collapsed taxa grey.
The output is a translation table of phylum to color, and a `ggplot2` plot.
```{r, include = T, fig.align = "center", fig.width = 8}
fantaxtic_bar(ps_tmp, color_by = "Phylum", label_by = "Species", other_label = "Other")
```

You can also color by family. As multiple OTUs/ASVs with the same name exist, you should make sure to generate unique labels.
```{r, include = T, fig.align = "center", fig.width = 8}
fantaxtic_bar(ps_tmp, color_by = "Phylum", label_by = "Family", other_label = "Other",
               gen_uniq_lbls = TRUE)
```

Perhaps a different sample ordering is more meaningful. Let's order the samples by the abundance of the collapsed category, labeled as "Other".
```{r, include = T, fig.align = "center", fig.width = 8}
fantaxtic_bar(ps_tmp, color_by = "Phylum", label_by = "Family", other_label = "Other",
               order_alg = "other.abnd")
```

Does the story change when we facet the plot by sample type? Add a faceting factor.
```{r, include = T, fig.align = "center", fig.width = 8}
fantaxtic_bar(ps_tmp, color_by = "Phylum", label_by = "Family", facet_by = "SampleType", other_label = "Other")
```
